### 2018-11-04 pilot_manuscript.Rmd
takin a Brain Break!

>>>>

To this point, most text analysis utilizing the Moral Foundations Dictionary operationalizes endorsement of any one of the foundations as percent occurence of words from the foundation's respective word list. As such, most analyses assume that zero percent occurence is indicative of no endorsement, while any non-zero percent occurence indicates endorsement of the foundation. This may not be sufficient in describing the true nature of the writer or speaker's endorsement of one of the sets of moral intuitions. A quick glance at the MFD words for *harm/care* reveals the presence of words that are more closely associated with universally accepted conceptions of *harm* over *care* and vice-versa [@Graham2009]. For example, the word "cruel" has relatively negative connotations compared to "benefit." For the *harm/care* foundation, it is conceivable that use of the word "cruel" might indicate a greater attentional focus of the idea of *harm* rather than *care*.

For *harm/care*, the definition of the foundation as well as its name clearly distinguishes between two somewhat opposite sides of an attentional continuum, with *harm* on the negative end and *care* on the positive side. In other words, the entries in the MFD for *harm/care* have somewhat clear positive and negative valences. The same pattern can be seen in the MFD entries for the other four foundations. *Purity/sanctity* features words that likely have a negative valence to most observers, including "disease"" and "trash," along with more positive words, including "right" and "sacred" [@Graham2009]. This, however, brings up other questions regarding the definition and names of the other four foundations apart from *harm/care*: *fairness/reciprocity*, *ingroup/loyalty*, *authority/respect*, and *purity/sanctity*. The latter four foudnations have names that are harder to understand as a valence continuum, as the concepts in the names are more similar, even to the point of being virtually synonymous in the case of *fairness/reciprocity*.

When considering the issue of positive versus negative valence in MFD words, the question of how texts are analyzed vis-a-vis the MFD remains. How can raw percentage of MFD word occurence capture the valence and focus of the writer or speaker? If 2% of a politician's speech features positive words (i.e. "benefit" and "defend") from the MFD *harm/care* list, how can researchers be sure the level and nature of the speaker's "endorsement" of the foundation equals that of another politician whose speech contained negatively connoted MFD words from the *harm/care* list? They would have equal endorsements as far as the numbers are concerned, but the words used and focus given are on opposite sides of the *harm/care* spectrum.

This issue is compounded by the fact the Moral Foundations Questionnaire and its subscales assume endorsement lies on a continuum.

##Valence
